MSc Data Science Project: Can a Convolutional Neural Network judge a book by its cover?

### 2020-08-11 Tex/MSc Data Science Project.tex
Write results and conclusion section

\subsection{Related Works} 
\label{sub:Related Works} 
Discuss the work of Iwana as the main source, reference their methods and results but keep much for later where relevant for comparison with my work.
% subsection Related Works end 
% section CNN Theory and Architectures
\section{Theory} 
\label{sec:Theory} 
\subsection{Components of Convolutional Neural Networks} 
\label{sub:Components_of_Convolutional_Neural_Networks} 
Describe the maths and the theory behind the various components of CNNs to the best of my ability, referring to the source papers for more complex combined units where sensible. 
% subsection Components_of_Convolutional_Neural_Networks end 
\subsection{Transfer Learning} 
\label{sub:Transfer_Learning} 
Describe the theory and maths behind transfer learning to the reader.
% subsection Transfer_Learning end

@@

Once the images had been processed we used a MobileNetV2 model on each processed training and validation dataset for 30 epochs with an early stopping patience of 15 (i.e. if no improvement was seen in validation accuracy over 15 epochs then we did not continue) and the best model for validation accuracy was kept. The model was trained using the default TensorFlow 2 settings of the adam optimiser with a loss function of sparse categorical cross entropy. More details of the set up are detailed in \cref{sec:Model_Training} as other than the epoch and early stopping the same configuration was used for the main training process. The test results of this model trained on each of the 3 datasets is presented in \cref{tab:preprocres} and the second method, padding the image with a white border, achieved the best result on the test data of 15.81\% accuracy.

@@

Iwana et al trained their models for 30,000 and 450,000 epochs for LeNet and AlexNet respectively, substantially more than we trained our models for, so it is likely that the results they achieved are better than they would have been if they had trained for only a few hundred epochs. It is possible, but not certain, that by training our models for a factor of 10 or 100 more epochs, and removing the early stopping criteria, that we could have achieved even better validation performance.
% subsection Training_Performance end

@@

\section{Evaluation} 
\label{sec:Evaluation_and_Further_Exploration} 
\subsection{Results} 
\label{sub:Results} 
Present and discuss the basic results for each model, including a comparison with Iwana with the caveat of a slightly different dataset and different pre-processing.

% section Results end 
\subsection{Further Analysis} 
\label{sub:Further_Analysis} 
Consider further, in depth, analysis on the best performing model, including some of the insight provided by Iwana.

Idea: compare class average colour by predicted and actual and compare to accuracy?
% section Further_Analysis end 
\subsection{Feature Visualisation} 
\label{sub:Feature_Visualisation} 
Detail the approach taken to try and visualise the ideal input to trigger the most accurate classes in an attempt to understand what feature the model is picking up on. Show output and discuss impact.
% section Feature_Visualisation end

\subsection{Discussion} 
\label{sub:Discussion} 
Discussion around the reasons we have seen some of these results if not already covered in the previous subsections.

Future work:
use different interpolation method
% subsection Discussion end 
% section Evaluation_and_Further_Exploration end

\section{Conclusion} 
\label{sec:Conclusion} 
Discuss the project as a whole, the impact of the results when compared to the literature, and what future work there could be off the back of this.
% section Conclusion end

>>>>

In this work we will use the terms \emph{class, genre, classification,} and \emph{category} interchangeably to describe the genre that a particular book cover belongs to.
\subsection{Related Works} 
\label{sub:Related Works} 
Discuss the work of Iwana as the main source, reference their methods and results but keep much for later where relevant for comparison with my work.
% subsection Related Works end 
% section CNN Theory and Architectures
\section{Theory} 
\label{sec:Theory} 
\subsection{Components of Convolutional Neural Networks} 
\label{sub:Components_of_Convolutional_Neural_Networks} 
Describe the maths and the theory behind the various components of CNNs to the best of my ability, referring to the source papers for more complex combined units where sensible. 
% subsection Components_of_Convolutional_Neural_Networks end 
\subsection{Transfer Learning} 
\label{sub:Transfer_Learning} 
Describe the theory and maths behind transfer learning to the reader.
% subsection Transfer_Learning end

@@

Once the images had been processed we used a MobileNetV2 model on each processed training and validation dataset for 30 epochs with an early stopping patience of 15 (i.e. if no improvement was seen in validation accuracy over 15 epochs then we did not continue) and the best model for validation accuracy was kept. The model was trained using the default TensorFlow 2 settings of the adam optimiser with a loss function of sparse categorical cross entropy. More details of the set up are detailed in \cref{sec:Model_Training} as other than the epoch and early stopping the same configuration was used for the main training process. The progress on the training and validation data can be seen in \cref{fig:prep_perf} and the test results of this model trained on each of the 3 datasets is presented in \cref{tab:preprocres}. The second method, padding the image with a white border, achieved the best result on the test data of 15.81\% accuracy.

\begin{figure}
	\centering
	\captionsetup{justification=centering}
	\includegraphics[scale=0.5]{prep_training.png}
	\caption{Performance of different pre-processing methods throughout training of MobileNetV2}
	\label{fig:prep_perf}
\end{figure}

@@

Iwana et al trained their models for 30,000 and 450,000 epochs for LeNet and AlexNet respectively, substantially more than we trained our models for, so it is likely that the results they achieved are better than they would have been if they had trained for only a few hundred epochs. It is possible, but not certain, that by training our models for a factor of 10 or 100 more epochs, and removing the early stopping criteria, that we could have achieved even better validation performance.

MobileNetV2 achieved the best validation accuracy early on in the training, where Inception-ResNeXt50 continued to see improvement until just before the 50th epoch/ ResNeXt50 sees its best performance coming after the 75th epoch, but as we will see this is not indicative of better performance the accuracy at any time never improved beyond the lowest accuracy for MobileNetV2 on the training or validation data despite being a more complex model and training for more epochs.
% subsection Training_Performance end

@@

\section{Evaluation} 
\label{sec:Evaluation_and_Further_Exploration} 
\subsection{Results} 
\label{sub:Results} 
Full results for all 3 models evaluated on the test set are available in \cref{tab:results_acc}, along with those reported by Iwana for ease of comparison. A visual representation of this for our models only is shown in \cref{fig:test_perf}. As a reminder, the dataset used for our models was slightly altered from the original set, and due to the use of a validation dataset did not contain the set of records; but for a comparison it is suitable. Inception-ResnetV2 is comparable to the results of AlexNet but is slightly more accurate when Top-3 predictions are taken into account. MobileNetV2 is close to LeNet in performance, with just slightly improvements in many genres. ResNeXt50 shows very odd behaviour, with many classes showing near 0\% accuracy even at Top-3, whilst some classes have a very high accuracy such as Biographies and Memoirs. This is likely an artefact of it overlapping specific classes which along with is discussed further in \cref{sub:Further_Analysis}.

At a per-genre level Inception-ResnetV2 shows the best results on 16 out of the 30 genres for top-1 predictions, followed by 10 for AlexNet. If we further explore the seemingly best performing classes in ResNeXt50 we can see by looking at the precision as presented in \cref{tab:test_prec} that in the genres where this model seems to perform well the precision is low (when compared to for example Inception-ResnetV2) which provides further evidence that for some classes it combines multiple genres and just predicts one of them i.e. whilst the model is able to correctly predict these \emph{chosen}	classes when it sees them, it is highly likely to predict other genres as one of these as well.

Overall the results from Inception-ResnetV2 are promising given the low training time and adjusted dataset; the potential improvements are discussed further in \cref{sub:Discussion}.

\begin{table}[]
	\centering
	\resizebox{\textwidth}{!}{%
	\begin{tabular}{lrrrrrrrrrrrrrr}
								   & \multicolumn{2}{c}{\textbf{LeNet*}}                                     & \multicolumn{1}{c}{\textbf{}} & \multicolumn{2}{c}{\textbf{AlexNet*}}                                   & \multicolumn{1}{c}{\textbf{}} & \multicolumn{2}{c}{\textbf{MobileNetV2}}                                & \multicolumn{1}{c}{\textbf{}} & \multicolumn{2}{c}{\textbf{Inception-ResnetV2}}                         & \multicolumn{1}{c}{\textbf{}} & \multicolumn{2}{c}{\textbf{ResNeXt50}}                                  \\
	\textbf{Genre}                 & \multicolumn{1}{c}{\textbf{Top 1}} & \multicolumn{1}{c}{\textbf{Top 3}} & \multicolumn{1}{c}{\textbf{}} & \multicolumn{1}{c}{\textbf{Top 1}} & \multicolumn{1}{c}{\textbf{Top 3}} & \multicolumn{1}{c}{\textbf{}} & \multicolumn{1}{c}{\textbf{Top 1}} & \multicolumn{1}{c}{\textbf{Top 3}} & \multicolumn{1}{c}{\textbf{}} & \multicolumn{1}{c}{\textbf{Top 1}} & \multicolumn{1}{c}{\textbf{Top 3}} & \multicolumn{1}{c}{\textbf{}} & \multicolumn{1}{c}{\textbf{Top 1}} & \multicolumn{1}{c}{\textbf{Top 3}} \\ \hline
	Arts \& Photography            & 5.8                                & 11.6                               &                               & 12.1                               & 31.1                               &                               & 11.1                               & 27.9                               &                               & \textit{13.2}                      & \textit{33.7}                      &                               & 3.7                                & 11.1                               \\
	Biographies \& Memoirs         & 5.3                                & 18.4                               &                               & 13.2                               & 29.5                               &                               & 10.0                               & 30.5                               &                               & \textit{23.2}                      & 46.3                               &                               & 20.0                               & \textit{60.0}                      \\
	Business \& Money              & 10.0                               & 25.3                               &                               & 12.6                               & 25.8                               &                               & 4.7                                & 18.4                               &                               & \textit{13.2}                      & \textit{33.7}                      &                               & 1.1                                & 3.2                                \\
	Calendars                      & 18.9                               & 37.9                               &                               & \textit{47.9}                      & \textit{65.3}                      &                               & 13.7                               & 25.8                               &                               & 25.8                               & 37.4                               &                               & 18.9                               & 24.7                               \\
	Children’s                     & 24.7                               & 42.1                               &                               & \textit{42.1}                      & \textit{61.6}                      &                               & 30.0                               & 52.1                               &                               & 18.9                               & 43.2                               &                               & 11.6                               & 34.7                               \\
	Comics \& Graphic Novels       & 15.8                               & 33.7                               &                               & 47.4                               & 67.9                               &                               & 40.5                               & 56.8                               &                               & \textit{56.3}                      & \textit{72.6}                      &                               & 47.9                               & 60.5                               \\
	Computers \& Technology        & 29.5                               & 42.8                               &                               & \textit{44.7}                      & \textit{59.5}                      &                               & 28.4                               & 43.7                               &                               & 41.1                               & 57.4                               &                               & 17.9                               & 26.8                               \\
	Cookbooks, Food \& Wine        & 14.2                               & 32.6                               &                               & \textit{43.7}                      & \textit{57.4}                      &                               & 23.7                               & 42.1                               &                               & 40.0                               & 55.8                               &                               & 7.9                                & 23.2                               \\
	Crafts, Hobbies \& Home        & 7.4                                & 22.1                               &                               & 17.4                               & 36.8                               &                               & 13.7                               & 30.5                               &                               & \textit{31.6}                      & \textit{60.0}                      &                               & 2.6                                & 15.3                               \\
	Christian Books \& Bibles      & \textit{8.4}                       & 23.7                               &                               & 7.4                                & \textit{26.3}                      &                               & 2.6                                & 12.1                               &                               & 4.2                                & 20.5                               &                               & 0.0                                & 0.5                                \\
	Engineering \& Transportation  & 10.0                               & 21.1                               &                               & 20.0                               & 34.7                               &                               & 15.3                               & 32.1                               &                               & \textit{20.0}                      & \textit{35.3}                      &                               & 6.8                                & 24.7                               \\
	Health, Fitness \& Dieting     & 4.2                                & 15.8                               &                               & 12.6                               & 29.5                               &                               & 4.2                                & 17.9                               &                               & \textit{12.6}                      & \textit{37.4}                      &                               & 3.7                                & 25.8                               \\
	History                        & 6.3                                & 16.8                               &                               & 12.6                               & 27.9                               &                               & 6.8                                & 21.6                               &                               & \textit{15.8}                      & \textit{38.4}                      &                               & 1.1                                & 2.1                                \\
	Humor \& Entertainment         & 5.3                                & 16.3                               &                               & 10.5                               & 22.6                               &                               & 0.0                                & 7.4                                &                               & \textit{16.3}                      & \textit{43.7}                      &                               & 0.0                                & 2.6                                \\
	Law                            & 14.7                               & 25.8                               &                               & 25.3                               & 38.4                               &                               & 16.3                               & 27.4                               &                               & 17.9                               & 37.9                               &                               & \textit{26.8}                      & \textit{45.3}                      \\
	Literature \& Fiction          & 3.2                                & 12.1                               &                               & \textit{11.1}                      & 22.6                               &                               & 6.3                                & 24.7                               &                               & 10.5                               & \textit{38.9}                      &                               & 0.5                                & 2.6                                \\
	Medical Books                  & 12.6                               & 30.0                               &                               & 19.5                               & 36.8                               &                               & 11.1                               & 27.4                               &                               & \textit{20.0}                      & \textit{45.8}                      &                               & 13.7                               & 38.9                               \\
	Mystery, Thriller \& Suspense  & 23.7                               & 40.0                               &                               & 34.2                               & 48.9                               &                               & 12.1                               & 28.9                               &                               & \textit{34.7}                      & \textit{59.5}                      &                               & 19.5                               & 28.9                               \\
	Parenting \& Relationships     & 14.7                               & 35.3                               &                               & \textit{24.2}                      & 39.5                               &                               & 24.2                               & \textit{43.7}                      &                               & 22.6                               & 42.1                               &                               & 4.7                                & 22.6                               \\
	Politics \& Social Sciences    & 3.7                                & 18.4                               &                               & 6.8                                & 21.6                               &                               & \textit{26.3}                      & \textit{55.8}                      &                               & 2.6                                & 13.2                               &                               & 0.5                                & 12.6                               \\
	Reference                      & 13.2                               & 26.8                               &                               & \textit{20.0}                      & \textit{34.2}                      &                               & 11.1                               & 21.6                               &                               & 8.9                                & 27.9                               &                               & 2.6                                & 18.4                               \\
	Religion \& Spirituality       & 8.4                                & 27.9                               &                               & 16.3                               & 31.6                               &                               & 7.9                                & 26.3                               &                               & \textit{36.3}                      & \textit{55.8}                      &                               & 0.0                                & 0.5                                \\
	Romance                        & 27.4                               & 43.2                               &                               & \textit{45.3}                      & \textit{60.5}                      &                               & 38.4                               & 52.1                               &                               & 40.5                               & 59.5                               &                               & 25.8                               & 52.1                               \\
	Science \& Math                & 8.4                                & 26.3                               &                               & 14.2                               & 29.5                               &                               & 3.7                                & 16.3                               &                               & \textit{26.3}                      & \textit{56.3}                      &                               & 0.0                                & 0.5                                \\
	Science Fiction \& Fantasy     & 14.7                               & 33.2                               &                               & \textit{35.8}                      & 52.6                               &                               & 21.6                               & 42.1                               &                               & 33.7                               & \textit{55.8}                      &                               & 18.9                               & 39.5                               \\
	Self-Help                      & 13.7                               & 31.6                               &                               & 14.2                               & 33.2                               &                               & 3.7                                & 17.9                               &                               & 11.6                               & 28.4                               &                               & \textit{18.9}                      & \textit{46.8}                      \\
	Sports \& Outdoors             & 5.3                                & 16.8                               &                               & 14.7                               & 28.4                               &                               & 1.1                                & 4.7                                &                               & \textit{27.4}                      & \textit{53.2}                      &                               & 0.0                                & 2.1                                \\
	Teen \& Young Adult            & 7.9                                & 17.4                               &                               & 12.1                               & 28.4                               &                               & 7.4                                & 28.9                               &                               & \textit{13.2}                      & \textit{35.3}                      &                               & 0.0                                & 0.5                                \\
	Test Preparation               & 47.9                               & 56.8                               &                               & \textit{68.9}                      & \textit{78.4}                      &                               & 34.2                               & 47.9                               &                               & 56.8                               & 74.2                               &                               & 50.5                               & 66.8                               \\
	Travel                         & 19.5                               & 33.7                               &                               & 33.2                               & 48.4                               &                               & 7.4                                & 14.2                               &                               & \textit{44.7}                      & \textit{64.2}                      &                               & 24.2                               & 41.1                               \\ \hline
	\textbf{Total Accuracy}         & \textbf{13.5}                      & \textbf{27.8}                      & \textbf{}                     & \textbf{24.7}                      & \textbf{40.3}                      & \textbf{}                     & \textbf{14.6}                      & \textbf{30.0}                      &                               & \textit{\textbf{24.7}}             & \textit{\textbf{45.4}}             &                               & \textbf{11.7}                      & \textbf{24.5}                      \\ \hline
	\textit{Number of best genres} & \textit{1}                         & \textit{0}                         & \textit{}                     & \textit{10}                        & \textit{8}                         & \textit{}                     & \textit{1}                         & \textit{2}                         & \textit{}                     & \textit{16}                        & \textit{17}                        & \textit{}                     & \textit{2}                         & \textit{3}                        
	\end{tabular}%
	}
	\caption{Top-1 and -3 Per-class recall (and overall accuracy) for the 3 trained models and the results presented by Iwana for comparison. Italics represent the best result for that genre in Top-1/3 recall/accuracy. \\
	\tiny{*Iwana results are based on a slightly different training and test set}
	}
	\label{tab:results_acc}
	\end{table}

\begin{figure}
	\centering
	\captionsetup{justification=centering}
	\includegraphics[scale=0.5]{accuracy_by_model.png}
	\caption{Top-1/3/5 recall (accuracy for all) of each model by genre}
	\label{fig:test_perf}
\end{figure}

\begin{table}[]
	\centering
	\resizebox{\textwidth}{!}{%
	\begin{tabular}{lrrr}
	\textbf{Predicted Genre}      & \multicolumn{1}{c}{\textbf{MobileNetV2}} & \multicolumn{1}{c}{\textbf{InceptionResnet-V2}} & \multicolumn{1}{c}{\textbf{ResNeXt50}} \\ \hline
	Arts \& Photography           & 17.8                                     & 20.8                                            & 25.9                                   \\
	Biographies \& Memoirs        & 11.5                                     & 19.3                                            & 5.5                                    \\
	Business \& Money             & 11.0                                     & 13.4                                            & 20.0                                   \\
	Calendars                     & 32.5                                     & 31.2                                            & 16.8                                   \\
	Children's Books              & 24.2                                     & 36.7                                            & 13.2                                   \\
	Christian Books \& Bibles     & 5.6                                      & 17.4                                            &                                        \\
	Comics \& Graphic Novels      & 19.7                                     & 46.5                                            & 17.1                                   \\
	Computers \& Technology       & 26.3                                     & 31.1                                            & 19.3                                   \\
	Cookbooks, Food \& Wine       & 19.1                                     & 65.0                                            & 16.3                                   \\
	Crafts, Hobbies \& Home       & 14.2                                     & 27.5                                            & 31.3                                   \\
	Engineering \& Transportation & 10.2                                     & 54.3                                            & 12.3                                   \\
	Health, Fitness \& Dieting    & 5.9                                      & 15.2                                            & 5.9                                    \\
	History                       & 13.5                                     & 20.3                                            & 66.7                                   \\
	Humor \& Entertainment        & 0.0                                      & 13.4                                            & 0.0                                    \\
	Law                           & 16.1                                     & 24.1                                            & 9.4                                    \\
	Literature \& Fiction         & 8.7                                      & 11.6                                            & 50.0                                   \\
	Medical Books                 & 12.4                                     & 16.5                                            & 12.5                                   \\
	Mystery, Thriller \& Suspense & 25.8                                     & 23.8                                            & 19.5                                   \\
	Parenting \& Relationships    & 9.8                                      & 33.9                                            & 8.2                                    \\
	Politics \& Social Sciences   & 5.5                                      & 11.1                                            & 5.3                                    \\
	Reference                     & 19.4                                     & 26.2                                            & 10.2                                   \\
	Religion \& Spirituality      & 9.6                                      & 16.7                                            &                                        \\
	Romance                       & 14.6                                     & 44.0                                            & 14.0                                   \\
	Science \& Math               & 11.9                                     & 11.8                                            & 0.0                                    \\
	Science Fiction \& Fantasy    & 18.3                                     & 39.5                                            & 13.6                                   \\
	Self Help                     & 12.3                                     & 22.9                                            & 7.2                                    \\
	Sports \& Outdoors            & 22.2                                     & 26.0                                            & 0.0                                    \\
	Teen \& Young Adult           & 9.5                                      & 14.2                                            & 0.0                                    \\
	Test Preparation              & 50.4                                     & 43.7                                            & 13.0                                   \\
	Travel                        & 50.0                                     & 17.3                                            & 8.3                                    \\ \hline
	\textbf{Total Precision}      & \textbf{14.6}                            & \textbf{24.7}                                   & \textbf{11.7}                          \\ \hline
	\end{tabular}%
	}
	\caption{Top-1 Per class and total precision for each model}
	\label{tab:test_prec}
	\end{table}

% section Results end 
\subsection{Further Analysis} 
\label{sub:Further_Analysis} 
In the paper by Iwana they discuss the potential reasons for the disparity between the recall by genre and investigate some examples. They call out consistency of colour, objects, and text styles are key elements that impact the ability for a genre to be able to be predicted well by their models. Here we focus on the colour component as the strongest driver for predictability and investigate the ways in which the models may be driven by colour of the cover images.

First, as displayed in \cref{fig:col_distributions}, we can see that whilst there is a wide range of average RGB values and width/aspect ratio (height is almost entirely consistent across genres), there are no standout values in colour alone that allow for easy genre classification. Width/aspect ratio, something that wasn't directly included in the model (but was indirectly by the use of the padding applied during the pre-processing), contains some clearer distinctions; in particular Calendars have a very different distribution of aspect ratios compared to other genres, and children books contains some of the more wider books than many other categories. Whilst none of this alone would be enough to provide good classification, it does suggest that explicit inclusion of dimensional information in a model could improve the performance when combined with the existing output.

\begin{figure}
	\centering
	\captionsetup{justification=centering}
	\includegraphics[scale=0.45]{distributions_col_ratio.png}
	\caption{Distributions of various measures by genre}
	\label{fig:col_distributions}
\end{figure}

The average overall colour per genre is presented in \cref{fig:avg_col} which does suggest some slight differences between the genres; similar to the callouts made by Iwana, Sci-Fi and Fantasy seem to be darker when compared to something like Cookbooks, Food and Wine. We can compare the average colour of the true classes vs those predicted by each model and visualise how these correlate with the recall per class as shown in \cref{fig:col_distance}. Here the colour distance is defined as

\begin{equation}
	d_{colour} = \sqrt{(R_2 - R_1)^2 + ((G_2 - G_1)^2 +((B_2 - B_1)^2 }
\end{equation}
where R/G/B are the average red, green, and blue values for each predicted and class respectively. We can see a small trends here, particular in MobileNetV2, where the closer the average colour of the true class the better the recall of that class. This is of course a somewhat self-fulfilling prophecy; the more you correctly predict the class the more the dataset will be the true class, however is highlights that even by just guessing based on closeness to average colour you can make some small improvements. This supports the claims made by Iwana in their work.

\begin{figure}
	\centering
	\captionsetup{justification=centering}
	\includegraphics[scale=0.45]{avg_col_plot.png}
	\caption{Average cover colour by genre}
	\label{fig:avg_col}
\end{figure}

\begin{figure}
	\centering
	\captionsetup{justification=centering}
	\includegraphics[scale=0.6]{col_dist_plot.png}
	\caption{Colour Distance vs Recall per genre by model}
	\label{fig:col_distance}
\end{figure}

Finally, we can review the \emph{class overlap} for each model. We define class overlap between class A and B for a single prediction as the ratio of the probability of prediction for class A and class B (where A has a higher probability) i.e.
\begin{equation}
	CO(A, B) \stackrel{\mathrm{def}}{=} \min\left\{\frac{P(A)}{P(B)}, \frac{P(B)}{P(A)}\right\}
\end{equation}
The per genre pair class overlap is defined as the average class overlap across all predictions. As we can see in \cref{fig:class_overlap}, where only class overlaps of greater than 0.5 are displayed, the ResNeXt50 model sees large amounts of class overlap between many classes, which agrees with what we saw before that certain classes were being \emph{confused} for each other by the model. Comparatively, MobileNetV2 sees only a few class overlaps greater than 0.5 and Inception-ResnetV2 sees none. When we examine those with higher overlap in MobileNetV2 we can begin to understand the potential true overlap between classes; Self-Help has overlap with Health, Fitness and Dieting, Science and Maths has overlap with Medical Books, and Religion and Spirituality has overlap with Christian Books and Bibles. All of these are logical overlaps and were a human tasked with classifying books based on their cover into these categories it is highly probable that different people would classify them differently.

\begin{figure}
	\centering
	\captionsetup{justification=centering}
	\includegraphics[scale=0.6]{class_pred_overlap.png}
	\caption{Average class overlap by model}
	\label{fig:class_overlap}
\end{figure}
% section Further_Analysis end 
\subsection{Feature Visualisation} 
\label{sub:Feature_Visualisation} 
The goal of using feature visualisation was to attempt to understand what, if any, underlying structure or information the models were using to aid their classification. To do this we use activation maximisation as implemented into the python package \emph{keras-vis}\cite{keras-vis} to do all the work. To produce the images the final layer activation function was changed from softmax to linear to ensure that class is being visualised as opposed to minimising other classes, and the loss function is simply defined as that linear activation (note here that loss is really a misnomer as we are trying to maximise this value as opposed to minimise it as we traditionally would) as discussed in \cref{sub:Activation_Maximisation}. L2 normalisation was used as the default to attempt to avoid pure static images being produced.

We ran the activation maximisation on the Inception-ResnetV2 model for the 3 best predicted genres (Test Prep, Comics \& Graphic Novels, and Travel) as we believed these had the best probability of returning useful images. The inputs to the model were then trained for 1000 epochs, at which point issues were encountered with the RAM on the machine. During these steps there was some improvement to the loss function however did not appear to converge and when compared to the progress seen on a pure imagenet model these saw very little movement and remained negative the entire time (where as the examples provided by the package are always positive after the first few epochs).

The results of this attempt can be seen in \cref{fig:act_max_output} and whilst it may be possible to argue that there are some differences between these images that you could prescribe to the common layout of these type of genres covers, it would be difficult to say this with any conviction. Overall the images are generic and provided no additional insight to understand what features the model is identifying to use for the classification.

\begin{figure}
	\centering
	\captionsetup{justification=centering}
	\includegraphics[scale=0.6]{visualize-dense-layer.png}
	\caption{Activation maximisation outputs for the 3 best predicted classes of the Inception-ResnetV2 model}
	\label{fig:act_max_output}
\end{figure}
% section Feature_Visualisation end

\subsection{Discussion \& Future Work} 
\label{sub:Discussion} 
Overall the Inception-ResnetV2 model provided the best results of the 3 models we tried, however it is far more complex than AlexNet but did not provide greatly improved results. The counterpoint to this is that the AlexNet model as trained by Iwana et al was trained for 450,000 epochs where out model was trained for less than 50 epochs total. We do not know the progress of the training of the model by Iwana but it would be reasonable to suggest that if their model had trained for less epochs, or ours for more, then there would be a greatly difference between the results. Despite training on slightly less data (due to the validation set), we can see that more modern and powerful CNNs are able to predict book genres with slightly improved accuracy compared to older methods.

There are lots of directions this work could continue to go in the future; as mentioned before there is still more investigation that can be done surrounding the pre-processing of data, either varying the interpolation methods for downsizing or re-testing the pre-processing methods used specifically on the end model architecture for a longer period of time. Our models could be trained for more epochs with a greater requirement on early stopping to allow for long term improvements to be made that may not be evident over a few hundred epochs. Also, additional information such as the width and height or just aspect ratio could be passed to the model as well in an attempt to improve accuracy this way. The final fully connected part of the network could also be increased in complexity to contain more layers as opposed to just the single layer used in our work.

If resources were no constraint some of the larger improvements to the work would be to tune the hyperparameters of the model, and even train the actual core part of the architecture itself rather than relying entirely on transfer learning and freezing the weights for this component. To do this would require a far greater dataset of images so a generator may have to be used or a larger dataset collected first. Finally, as shown by the results in \cref{tab:results_acc} there is not a single model that performs best across all genres, so an ensemble method may be able to increase the accuracy at the cost of increased complexity. Many of these suggestions are beyond the scope of many citizen data scientists resources and would also lead to a model that may have a prediction time that makes it prohibitive for any real-time use.
% subsection Discussion end 
% section Evaluation_and_Further_Exploration end

\section{Conclusion} 
\label{sec:Conclusion} 
In this work we have attempted to improve on the work by Iwana et al and better answer the question can a CNN judge a book by its cover. We reprocessed the dataset initially collected by Iwana and applied multiple pre-processing techniques in an attempt to identify the best method to deal with the impact of different shaped images. We then proceeded to train 3 modern models via transfer learning on a padded version of the dataset and evaluated the results. Finally we attempted to produce ideal class images using activation maximisation for the best model and genres.

We have seen that using a more modern CNN architecture, namely Inception-ResnetV2, we are able to improve on the results published by Iwana slightly whilst training for substantially less epochs. We have also provided some further supporting insight into the components that factor into book cover design i.e. colour, and identified where classifications are failing where this might be due to overlap between classes.

There is still lots to explore in this space as we discussed in \cref{sub:Discussion} and hopefully this work provides a springboard to enable more complex CNNs to be used for genre identification of book covers in the future. The work itself is unlikely to provide any value for real-world use as the accuracy is too low to be of use for either consumers or even cover designers as a form of QA. Genre prediction from cover images remains a complex topic and given the subjectivity of the classifications is unlikely to ever be fully solved, however the work produced here gets a little closer than we were before. So it seems for now that, if you are a CNN, you should follow the original advice and not try to judge a book by its cover.
% section Conclusion end
